Volunteers in the Clouds: A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2016-01-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Permission to cut this sections in SOA
Fully removed all material in the other paper
And additionally reduced content to 12 pages to comply with requirements.

The fulfillment of these properties could be achieved by using any of the
languages and associated technologies, mentioned above, provided that the user had
installed the required virtual machine plugin first. 
% (Paloma) These should be separate sentences
However, embedding a
distributed computation client in the browser has the advantage of
providing a single and universal way of accessing resources via an
URI, something that is not guaranteed by using desktop applications. 
Baratloo et al. tried in 1996 to create a framework with
the aforementioned qualities called {\em Charlotte}
\cite{baratloo1999charlotte}. This framework, which implemented a
distributed shared memory, was inspired by a previous work
on parallel computing libraries PVM and MPI and solved most of the issues by
using the Java virtual machine embedded in the browser,
which provided
security to the user, uniformity of computing environment
independently of the operating system, and access to a common data store in the
server. Additionally, since the Java virtual machine is a sandbox, it was safe to the user
and as it runs compiled bytecode it also guaranteed a certain level of
performance. % (Paloma) I've changed a bit this sentence to unify the use of past tense when talking about Charlotte
Other issues such as scheduling, which are related to adaptive 
parallelism, were solved on the server. It is interesting to note that,
even in the early stages of the web, the possibility of
% (Paloma) Re-write. For example, "but it is interesting to note that, even in the early stages of the web, the possibility of..."
browser-based distributed computing was already researched.

@@

Initial experiments were set up using the OpenShift
PaaS, which provides a free tier within which these runs could be performed. That way, and as required, the hardware
and cloud cost for this experiment were zero. Experiments were
announced through a post in Twitter and other social networks, and
results were published here \cite{DBLP:conf/gecco/GuervosG15}. For the
purpose of this paper, we repeated the announcement several times
through the month of April and then by the beginning of August. All
in all, we have the set of runs with the characteristics shown in
Table \ref{tab:runs}. In general, every experiment took several
days. No particular care was taken about the time of the announcement
or the particular wording. Every {\em experiment} consisted in running
until the solution of the 40-trap problem was found. When the correct
solution was sent to the server, the counter was updated and the pool
of solutions reset to the void set. There was no special intention to wait
until all clients had finished, thus it might happen that, in fact,
the islands running in the browser {\em spill} from one experiment to
the next. However, previous experiments have proven that the influence
of these islands in the next experiment is indeed negligible.
%
\begin{table*}
\caption{Characteristics of the volunteer computing initial runs. \label{tab:runs}}
\begin{center}
\begin{tabular}{l|rr}
\hline
Date & \# experiments & Different IPs \\
\hline
April 4th 4/4 & 57 & 191 \\
April 24th 4/24 &  231 & 559 \\
July 31th 7/31 & 97 & 179 \\
\hline
\end{tabular}
\end{center}
\end{table*}
%
\begin{table*}
\caption{Summary of time per run, number of IPs and number of {\tt
    PUT}s per IP in the initial runs. \label{tab:summary:os}}
\begin{center}
\begin{tabular}{l|ccccccc}
\hline
Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
\hline
4/4 & 5 & 16 & 2040 & 18 & 14.29\% & 5.36\% & 0.0082 \\
4/24 &  5 & 29 & 732 & 11 & 47.39\% & 3.91\% & 0.0934\\
7/31 & 5 & 14 & 260 & 23 & 27.08\% & 1.04\%  & 0.1741\\
\hline
\end{tabular}
\end{center}
\end{table*}
%
The table shows that every run included more than 50 experiments. The
number of different IPs intervening in them varied from more than one
hundred to more than five hundred in the second experiment.

>>>>

The fulfillment of these properties could be achieved by using any of the
languages and associated technologies mentioned above, provided that the user had
installed the required virtual machine plugin first. 
% (Paloma) These should be separate sentences
However, embedding a
distributed computation client in the browser has the advantage of
providing a single and universal way of accessing resources via an
URI, something that is not guaranteed by using desktop applications. 
Baratloo et al. tried in 1996 to create a framework with
the aforementioned qualities called {\em Charlotte}
\cite{baratloo1999charlotte}. This framework, which implemented a
distributed shared memory, was inspired by a previous work
on parallel computing libraries PVM and MPI and solved most of the issues by
using the Java virtual machine embedded in the browser,
which provided
security to the user, uniformity of computing environment
independently of the operating system, and access to a common data store in the
server. Additionally, since the Java virtual machine is a sandbox, it was safe to the user
and as it runs compiled bytecode it also guaranteed a certain level of
performance. % (Paloma) I've changed a bit this sentence to unify the use of past tense when talking about Charlotte
Other issues such as scheduling, which are related to adaptive 
parallelism, were solved on the server. It is interesting to note that,
even in the early stages of the web, the possibility of
% (Paloma) Re-write. For example, "but it is interesting to note that, even in the early stages of the web, the possibility of..."
browser-based distributed computing was already researched.

@@

Initial experiments were set up using the OpenShift
PaaS, which provides a free tier within which these runs could be performed. That way, and as required, the hardware
and cloud cost for this experiment were zero. Experiments were
announced through a post in Twitter and other social networks, and
results were published here \cite{DBLP:conf/gecco/GuervosG15}. For the
purpose of this paper, we repeated the announcement several times
through the month of April and then by the beginning of August. All
in all, we have the set of runs with the characteristics shown in
Table \ref{tab:summary:os}. In general, every experiment took several
days. No particular care was taken about the time of the announcement
or the particular wording. Every {\em experiment} consisted in running
until the solution of the 40-trap problem was found. When the correct
solution was sent to the server, the counter was updated and the pool
of solutions reset to the void set. There was no special intention to wait
until all clients had finished, thus it might happen that, in fact,
the islands running in the browser {\em spill} from one experiment to
the next. However, previous experiments have proven that the influence
of these islands in the next experiment is indeed negligible.
%
\begin{table*}
\caption{Summary of time per run, number of IPs and number of {\tt
    PUT}s per IP in the initial runs. \label{tab:summary:os}}
\begin{center}
\begin{tabular}{l|ccccccccc}
\hline
Date & Experiments &\#IPs &\#IPs &  \#IPs & Time (s) &  \#{\tt PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment\\
     & \# & Different & Median & Max & Median &  Median & & & correlation \\ 
\hline
4/4 & 57 & 191 & 5 & 16 & 2040 & 18 & 14.29\% & 5.36\% & 0.0082 \\
4/24 & 231 & 559 & 5 & 29 & 732 & 11 & 47.39\% & 3.91\% & 0.0934\\
7/31 & 97 & 179 & 5 & 14 & 260 & 23 & 27.08\% & 1.04\%  & 0.1741\\
\hline
\end{tabular}
\end{center}
\end{table*}
%
The table shows that every run included more than 50 experiments. The
number of different IPs intervening in them varied from more than one
hundred to more than five hundred in the second experiment.
